# Factors Explaining the Spatial Variations in the Results of the Swiss Popular Initiative ”Against Mass Immigration”
This code is sparsley commented - please refer to the report for more details and the documentation of tmap
https://cran.r-project.org/web/packages/tmap/vignettes/tmap-nutshell.html 

## Configure R and load data

```{r, echo=TRUE, eval=FALSE}
install.packages("spgwr")
```

```{r, echo=TRUE, eval=FALSE}
# For spatial data handling
library(rgdal)
library(car)
library(spdep)
library(RColorBrewer)
library(rgeos)
# For charting
library(tmap)
library(cartogram)
library(ggplot2)
library(gridExtra)
library(GGally)
# For data loding and munging
library(readr)
library(dplyr)
library(tidyr)
# For spatial stats
library(GWmodel)
library(spdep)
# For cluster analysis
library(cluster)
library(car)
library(knitr)
library(spgwr)
```


```{r, echo=TRUE, eval=FALSE}
# read .csv containing referendum results data into an R DataFrame. 
referendum_data<- read_csv("./data/referendum.csv")
census_data<- read_csv("./data/census.csv")

# iterate over OA level data and compute summary statistics on relevant variables 
# Read in shapefile 
ch_kanton <- readOGR(dsn = "shapefiles", layer = "Kantone")
ch_kanton@data$AK <- as.character(ch_kanton@data$AK)
# merge results and census data with SpatialDataFrame  
# inner_join function.
ch_kanton@data <- inner_join(ch_kanton@data, referendum_data,  by=c("AK" =  "AK"))
ch_kanton@data <- inner_join(ch_kanton@data, census_data, by=c("AK" =  "AK"))
# rename  
rm(referendum_data)
rm(census_data)
```

## Exploring spatial variation in the Yes:No vote 
We first observe how voting preference, the percentage of people voting against mass migration in CH cantons, varies geographically by plotting the raw scores on a map.

```{r, echo=TRUE, eval=FALSE}
# A SpatialDataFrame must always be supplied to tm_shape(). To tm_fill(), we identify the variable values 
# on which polygons should be coloured as well as information such as the colour mapping (sequential, diverging 
# or continuous) and palette to use. Many, many layout specifications are available in tm_layout. Type ?tm_layout
# into the Console for a complete list.
outcomemap <- tm_shape(ch_kanton) +  
  tm_fill(col="cross_border",style="cont",palette="BrBG", size=0.2, id="AK", title="") + 
  tm_layout(
     title="% points",
     title.snap.to.legend=TRUE,
     title.size=0.8,
     legend.text.size=0.6,
     title.position = c("left", "center"),
     legend.position = c("left","center"),
     frame=FALSE,
     legend.outside=TRUE)

outcomemap
```

### Cartogram 

```{r, echo=TRUE, eval=FALSE}
ch_kanton_carto <- cartogram(ch_kanton, "Electorate", itermax=5)
```

```{r, echo=TRUE, eval=FALSE}
# Plot the SpatialDataFrame via tm_shape() and tm_fill()
tm_shape(ch_kanton_carto) +  
  tm_fill(col="Yes_No",style="cont",palette="BrBG", size=0.2, id="AK", title="") + 
  tm_layout(
     title="% point against mass immigration",
     title.snap.to.legend=TRUE,
     title.size=0.8,
     legend.text.size=0.6,
     title.position = c("right", "center"),
     legend.position = c("right","center"),
     frame=FALSE,
     legend.outside=TRUE)

```


```{r, echo=TRUE, eval=FALSE}
# Interactive mode.
tmap_mode("view")
# Static mode:
tmap_mode("plot")
```


## What drives Canton voting preference?

To explore these variables' effect on voting preference, we can create a set of scatter plots that display relationships between each explanatory variable and the vote outcome. Note that in a similar way that spatial units in cartograms are sized according to a variable of interest, we size the points in the scatter plots by their electoral size and colour by Region.

```{r, echo=TRUE, eval=FALSE}
ch_kanton@data %>%
  ggplot(aes(x=private_transport_to_work, y=Yes_No))+ 
  geom_point(aes(fill=Yes_No, size=Electorate),colour="#525252",pch=21) +
  stat_smooth(method=lm, se=FALSE, size=1, colour="#525252")+
  scale_fill_distiller("BrBG", type="div", direction=1, guide="colourbar", limits=c(-0.29,0.29))+
  theme_bw()+
  theme(legend.position="none")+
  ggtitle(paste("correlation:",round(cor.test(ch_kanton@data$Yes_No,ch_kanton@data$foreign_nationals)$estimate,2)))

```

To create a scatter plot for each variable vs the Yes/No vote at once we can create a gird-plot via facet_wrap.

```{r, echo=TRUE, eval=FALSE}
ch_kanton@data %>%
  gather(c(population1000,age_65_plus,living_city,foreign_nationals,roman_catholic,evangelical,without_confession,unemployment,gdp_capita,change_gdp,primary_sector,secondary_sector,tertiary_sector,homeownership,cars,puplic_transport_to_work,private_transport_to_work,socialsecuritybeneficiary,degree_educ,criminal_offences_cc,blue_collar), key = "expl_var", value="canton_prop") %>%
  ggplot(aes(x=canton_prop, y=Yes_No))+ 
  geom_point(aes(fill=Yes_No, size=Electorate),colour="#525252",pch=21) +
  stat_smooth(method=lm, se=FALSE, size=1, colour="#525252")+
  scale_fill_distiller("BrBG", type="div", direction=1, guide="colourbar", limits=c(-0.29,0.29))+
  facet_wrap(~expl_var, scales="free")+
  theme_bw()+
  theme(legend.position="none")
  #ggtitle(paste("correlation:",round(cor.test(ch_kanton@data$Yes_No,ch_kanton@data$immigrants)$estimate,2)))
#cars,criminal_offences_cc,degree_educ,SVP,small_right_wing,private_transport_to_work,roman_catholic,age_65_plus
```

Let's narrow it down to the variables with the largest correlation.

```{r, echo=TRUE, eval=FALSE}
corrfacet_canton <- ch_kanton@data %>%
  gather(c(degree_educ,age_65_plus,private_transport_to_work,homeownership,foreign_nationals,unemployment,change_gdp,blue_collar,criminal_offences_cc,pop_change,cross_border), key = "expl_var", value="canton_prop") %>%
  ggplot(aes(x=canton_prop, y=Yes_No))+ 
  geom_point(aes(fill=Yes_No, size=Electorate),colour="#525252",pch=21) +
  stat_smooth(method=lm, se=FALSE, size=1, colour="#525252")+
  scale_fill_distiller("BrBG", type="div", direction=1, guide="colourbar", limits=c(-0.29,0.29))+
  facet_wrap(~expl_var, scales="free")+
  theme_bw()+
  theme(legend.position="none")

corrfacet_canton

```


```{r}
ggsave("corrfacet_canton.eps", plot = corrfacet_canton,width = 10)
```

Note that the Tessin is an outlier in almost all scatter plots.


## Basic Robust Regression GWR model selection

Bandwidth was choosen arbitrarly at 50 

```{r, echo=TRUE, eval=FALSE}}

DeVar <- "Yes"
InDeVars <- c("pop_change","pop_density","age_65_plus","living_city","foreign_nationals","roman_catholic","unemployment","homeownership","puplic_transport_to_work","private_transport_to_work","socialsecuritybeneficiary","degree_educ","criminal_offences_cc","blue_collar")
model.sel <- model.selection.gwr(DeVar, InDeVars, data = ch_kanton,kernel = "bisquare", adaptive = TRUE, bw = 50)
sorted.models <- model.sort.gwr(model.sel, numVars = length(InDeVars),ruler.vector = model.sel[[2]][,2])
model.list <- sorted.models[[1]]
model.view.gwr(DeVar, InDeVars, model.list = model.list)
plot(sorted.models[[2]][,2], col = "black", pch = 20, lty = 5,main = "GWR model selection procedure",ylab = "AICc", xlab = "Model number", type = "b")

```

```{r}
# AIC of best performing model
sorted.models[[2]][,2][84]
# coefficients selected 
sorted.models[[1]][[84]][[1]]
```

Calculate VIF scores to investigate potential collinearity: 

```{r}
lm.global <- lm(Yes~criminal_offences_cc+degree_educ+age_65_plus+private_transport_to_work+homeownership+foreign_nationals+unemployment+socialsecuritybeneficiary, data=ch_kanton@data)
summary(lm.global)
vif(lm.global)

```

Correlation Matrix:

```{r, echo=TRUE, eval=FALSE}
# Generate correlation coefficient matrix
ch_kanton@data %>% 
 select(Yes_No,pop_change,pop_density,age_65_plus,living_city,foreign_nationals,roman_catholic,unemployment,change_gdp,homeownership,puplic_transport_to_work,private_transport_to_work,socialsecuritybeneficiary,degree_educ,criminal_offences_cc,blue_collar) %>%
ggcorr(label=TRUE,nbreaks=5,  geom = "text", hjust = 1, size = 2, palette = "RdBu", label_alpha=0.1,label_size
=2)

corrmatrix <-ch_kanton@data %>% 
 select(Yes,degree_educ,age_65_plus,private_transport_to_work,homeownership,foreign_nationals,unemployment,change_gdp,blue_collar,criminal_offences_cc,pop_change) %>%
ggcorr(label=TRUE,nbreaks=5,  geom = "text", hjust = 1, size = 4, palette = "RdBu", label_alpha=0.1,label_size
=4,layout.exp = 2)
corrmatrix

```


```{r}
ggsave("corrmatrix.eps", plot = corrmatrix,width = 10)
```


Plot one of the correlation coefficients
```{r, echo=TRUE, eval=FALSE}
# Note that we use the 'optimal' bandwidth derived from the bw.gwr() function.
gw_ss <- gwss(ch_kanton, vars  =  c("Yes","degree_educ","age_65_plus","private_transport_to_work","homeownership","foreign_nationals","unemployment","blue_collar","criminal_offences_cc","cross_border","change_gdp"),
                  kernel = "bisquare", adaptive = TRUE, bw = 24, quantile = TRUE)

gwcorr <- tm_shape(gw_ss$SDF) +
  tm_fill(col=colnames(gw_ss$SDF@data[119:119]), title="", style="cont",palette="BrBG", size=0.2) + 
  tm_facets(free.scales = FALSE) +
  tm_layout(
     title=colnames(gw_ss$SDF@data[119:119]),
     frame=FALSE,
     title.snap.to.legend=TRUE,
     title.size=0.8,
     title.position = c("left", "top"),
     inner.margins = c(0,0,0.15,0),
     legend.title.size=1,
     legend.text.size=0.6,
     legend.outside=FALSE)
gwcorr
```


```{r}
save_tmap(gwcorr, "gwcorr_cross_border.png", height=5)

```


### Refined model (Model 8)
Due to high VIF scores and correlation coefficients, a sparse model with only two variables was selected: unemployment and private_transport_to_work 


```{r}
# Determine optimal bandwith for the refined GWR model:
bw.gwr.1 <- bw.gwr(Yes ~ private_transport_to_work + unemployment, data = ch_kanton,approach = "AICc", kernel = "bisquare", adaptive = TRUE)
bw.gwr.1
```

```{r}
# Create GW regression model
gwr.res <- gwr.basic(Yes ~ 
    + unemployment + private_transport_to_work, data = ch_kanton,bw = 23, kernel = "bisquare", adaptive = TRUE, F123.test = TRUE)
print(gwr.res)
```


```{r}
# Plot the residuals
gwr.res.resids <- tm_shape(gwr.res$SDF) +  
  tm_fill(col="residual",style="cont",palette="RdBu", size=0.2, id="AK", title="") + 
  tm_layout(
     title="Residuals",
     title.snap.to.legend=TRUE,
     title.size=0.8,
     legend.text.size=0.6,
     title.position = c("left", "top"),
     legend.position = c("left","top"),
     frame=FALSE,
     legend.outside=TRUE)

gwr.res.resids
```
```{r}
save_tmap(gwr.res.resids, "residual_map_2.png", height=5)
```


```{r}
# Global condition number
nobs <- dim(ch_kanton)[1]
lcrm1 <- gwr.lcr(Yes ~ private_transport_to_work+unemployment + cross_border, data = ch_kanton, bw = nobs,kernel = "boxcar", adaptive = TRUE)
summary(lcrm1$SDF$Local_CN)
```


### Adding addtional (external) variable cross_border - Model8+ 

```{r}
gwr_plus.res <- gwr.basic(Yes ~ 
    + unemployment + private_transport_to_work + cross_border, data = ch_kanton,bw = 23, kernel = "bisquare", adaptive = TRUE, F123.test = TRUE)
print(gwr_plus.res)
```

```{r}
# Plot residuals
gwr_plus.res.resids <- tm_shape(gwr.res$SDF) +  
  tm_fill(col="residual",style="cont",palette="RdBu", size=0.2, id="AK", title="") + 
  tm_layout(
     title="Residuals",
     title.snap.to.legend=TRUE,
     title.size=0.8,
     legend.text.size=0.6,
     title.position = c("left", "top"),
     legend.position = c("left","top"),
     frame=FALSE,
     legend.outside=TRUE)

gwr_plus.res.resids
```


## BKW condition numbers for various models
```{r}

test.CN <- function(model, data) {
lcrmx.bw <- bw.gwr.lcr(model, data = data, kernel = "bisquare",
adaptive = TRUE)
print(model)
print(lcrmx.bw)
lcrmx <- gwr.lcr(model, data = data, bw = lcrmx.bw,
kernel = "bisquare", adaptive = TRUE)
print(summary(lcrmx$SDF$Local_CN))
lcrmx$SDF$Local_CN
}
data <- ch_kanton

model1 <- as.formula(Yes~criminal_offences_cc+blue_collar+degree_educ+age_65_plus+change_gdp+homeownership+foreign_nationals+unemployment+private_transport_to_work)
Model1  <- test.CN(model1, data)


model2 <- as.formula(Yes~blue_collar+degree_educ+age_65_plus+change_gdp+homeownership+foreign_nationals+unemployment+private_transport_to_work)
Model2 <- test.CN(model2, data)

model3 <- as.formula(Yes~degree_educ+age_65_plus+change_gdp+homeownership+foreign_nationals+unemployment+private_transport_to_work)
Model3 <- test.CN(model3, data)

model4 <- as.formula(Yes~age_65_plus+change_gdp+homeownership+foreign_nationals+unemployment+private_transport_to_work)
Model4 <- test.CN(model4, data)

model5 <- as.formula(Yes ~ change_gdp+homeownership+foreign_nationals+unemployment+private_transport_to_work)
Model5 <- test.CN(model5, data)

model6 <- as.formula(Yes ~ homeownership+foreign_nationals+unemployment+private_transport_to_work)
Model6 <- test.CN(model6, data)

model7 <- as.formula(Yes ~ foreign_nationals+unemployment+private_transport_to_work)
Model7 <- test.CN(model7, data)

model8 <- as.formula(Yes ~ unemployment+private_transport_to_work)
Model8 <- test.CN(model8, data)

model8plus <- as.formula(Yes ~ unemployment+private_transport_to_work+cross_border)
Model8plus <- test.CN(model8plus, data)

```

```{r}

bkwbox <- boxplot(Model1,Model2,Model3,Model4,Model5,Model6,Model7,Model8,Model8plus, ylab = "BKW condition number", xlab = "",names = c("Model1","Model2","Model3","Model4","Model5","Model6","Model7","Model8","Model8+"))

```
